A Meta-Analysis of Music Emotion Recognition Studies

### 2025-03-28 manuscript/manuscript.qmd
Font edits to heatmap and MS edits to introduction

MER tasks commonly apply regression and classification techniques to diverse datasets with features drawn from both music (e.g., audio, MIDI, metadata) and participants (e.g., demographic information, survey responses, physiological signals, etc.). To facilitate model comparison, several databases have been shared publicly, including MediaEval [@soleymani2013], DEAM [@aljanaki2017developing], and AMG1608 [@chen2015amg1608]. These datasets represent Western pop music, are moderate in terms of the size (containing from 744 to 1802 music excerpts) and have been manually annotated by a relatively large number of participants (either by experts, students, or crowdsourced workers). Many datasets include audio features analyzed with software toolboxes for audio analysis such as OpenSMILE [@eyben2010opensmile] and MIRToolbox [@lartillot2007matlab] -- enabling models that predict ratings using tens, or even hundreds, of audio features.

REVISE:
Large datasets are ideal for generalizing predictions across diverse musical samples and are commonly employed in machine learning applications focused on maximizing accuracy of a MER system.

IDEA: LARGE data requirements for machine-learning/deep learning, but annotating music is resource-intensive.

>>>>

MER tasks commonly apply regression and classification techniques to diverse datasets with features drawn from both music (e.g., audio, MIDI, metadata) and participants (e.g., demographic information, survey responses, physiological signals, etc.). To facilitate model comparison, several databases have been shared publicly, including _MediaEval_ [@soleymani2013], _DEAM_ [@aljanaki2017developing], and _AMG1608_ [@chen2015amg1608]. These datasets represent Western pop music, are moderate in terms of the size (containing from 744 to 1802 music excerpts) and have been manually annotated by a relatively large number of participants (either by experts, students, or crowdsourced workers). Many datasets include audio features analyzed with software toolboxes for audio analysis such as _OpenSMILE_ [@eyben2010opensmile] and _MIRToolbox_ [@lartillot2007matlab] -- enabling models that predict ratings using tens, or even hundreds, of audio features.

Large datasets are needed to generalize predictions across diverse samples, especially for deep learning and complex machine-learning models that require extensive pre-training. However, the dataset sizes in MER tend to be modest in comparison, as direct annotation of music examples is resource-intensive. In contrast, datasets in the visual domain are often significantly larger (e.g., _EmoSet_ with 118,102 images and _AffectNet_ with 450,000, see @yang2023emoset;@mollahosseini2017affectnet). Some efforts have been made to scale up MER datasets by inferring emotions from tags (_MTG-Jamendo_ with 18,486 excerpts @bogdanov2019mtg and _Music4all_ with 109,269 excerpts @santana2020), but these have not found their way into standard emotion prediction or classification tasks yet.
